---
title: "HW 02"
author: "Joey Garcia"
format: 
  html:
    embed-resources: true
toc: true
---

```{r setup and load libraries}
if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse, lubridate, glue, scales, dplyr, ggthemes, # most necessary libs.
               ggridges, # density ridge plots and themes
               dsbox,    # data science in a box (datasets)
               fs,       # filesystem lib.
               janitor   # clean data 
               )       

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)

```

## 1 - A new day, a new plot, a new geom

```{r}
#| label: ridgeline-plot-setup

# Load and test data
airbnb <- dsbox::edibnb
glimpse(airbnb)
summary(airbnb)

```

```{r}
#| label: ridgeline-plot 

# review scores of Edinburgh neighborhoods ridgeline plot
airbnb |>
  filter(between(review_scores_rating, 85, 100)) |> # reduces axis
  filter(!is.na(neighbourhood)) |> # filters empty neighborhood rows
  ggplot(aes(x = review_scores_rating, y = neighbourhood, fill = neighbourhood)) +
  geom_density_ridges(scale = 4, rel_min_height = 0.1) + # scale and cut tail of density
  scale_fill_cyclical(values = c("skyblue1", "cornsilk4")) +
  scale_y_discrete(expand = c(0, 0)) + 
  scale_x_continuous(expand = c(0, 0)) + # reduces padding
  coord_cartesian(clip = "off") +
  labs(
    title = "Ridgeline plot", 
    subtitle = "listings from Airbnb in Edinburgh, Scotland",
    x = "Review Score Ratings",
    y = "Edinburgh\nNeighborhoods", 
    caption = "Source: From Opensource dataset dsbox known as 'edibnb'."
  )+
  theme_ridges()

```

## 2 - Foreign Connected PACs

```{r}
#| label: Foreign-connected-pac-data-cleaning
# get a list of files with "Foreign Connected PAC" in their names
list_of_files <- dir_ls(path = "data", regexp = "Foreign Connected PAC")

# read all files and row bind them
# track file name with column called year
pac <- read_csv(list_of_files, id = "year")

# Clean data
pac <- pac |>
  # Normalize names
  clean_names() |>
  
  # year into <int> 
  mutate(year = str_extract(year, "(?<=-)\\d{4}")) |>
  
  # separate columns by delimiter
  separate(country_of_origin_parent_company,
           into = c("country_of_origin", "parent_company"),
           sep = "/"
           ) |>
  
  # Change data type to <dbl>
  mutate(total = as.numeric(gsub("[$,]", "", total))) |>
  mutate(dems = as.numeric(gsub("[$,]", "", dems))) |>
  mutate(repubs = as.numeric(gsub("[$,]", "", repubs))) |>
  
  # pivot longer by "party"
  pivot_longer(cols = c("dems", "repubs"),
               names_to = "party",
               values_to = "amount"
               ) |>
  # rename columns
  mutate(party = recode(party, dems = "Democrat",repubs = "Republican"))

unique(pac$country_of_origin)
```


```{r}
#| label: Foreign-connected-pac-UK-contributions

pac |>
  filter(country_of_origin == "UK") |>
  # Group & summarize total contributions
  group_by(year, party) |>
  summarise(total_contributions = sum(amount), .groups = "drop") |>
  
  ggplot(aes(x = year, y = total_contributions, colour = party, group = party)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M")) +
  labs(
    title = "Contributions to US political parties from UK connected PACs",
    x = "Year",
    y = "Total Amount",
    colour = "Party"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(hjust = 0),
    axis.title.x = element_text(hjust = 0),
    legend.position = c(0.95, 0.05),
    legend.justification = c("right", "bottom")
  )
```

``` {r}
#| label: Foreign-connected-pac-Israel-contributions

pac |>
  filter(country_of_origin == "Spain") |>
  # Group & summarize total contributions
  group_by(year, party) |>
  summarise(total_contributions = sum(amount), .groups = "drop") |>
  
  ggplot(aes(x = year, y = total_contributions, colour = party, group = party)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "Contributions to US political parties from Spain connected PACs",
    x = "Year",
    y = "Total Amount",
    colour = "Party"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(hjust = 0),
    axis.title.x = element_text(hjust = 0),
    legend.position = c(0.95, 0.65),
    legend.justification = c("right", "bottom")
  )
```

### Foreign Connected PACs interpretation (Spain)

Spain connected PACs initially contributed the majority of the contributions to the Republican Party. During the 2020 US election year, the paradigm shifted and the Democrat Party received the majority of the contributions. As we move into 2022, we start to see more contributions going to the Democrat Party.  

## 3 - Median housing prices in the US

```{r}
#| label: lab

```

## 4 - Expect More. Plot More.

```{r}
#| label: lab

```

## 5 - Mirror, mirror on the wall, who’s the ugliest of them all?

```{r}
#| label: lab

```
